---
title: "SGupta_HW01"
format: pdf
editor: visual
---

## ST503 Assignment 1

## Running Code

##  (A) Write a brief description of the dataset. Produce some numerical and graphical summaries of the dataset.
```{r}
 #install.packages('faraway')
```

```{r} 
    library(faraway)
    data(teengamb)
    str(teengamb)
    print("Print Summary")
    summary(teengamb)
    print("Print top 5 values")
    head(teengamb)
```

## The dataset contains 47 observations of 5 variables:
##  sex: Binary variable (0 = Male, 1 = Female)
##  status: ranges from 18 to 75, with a mean of 45.23.
##  income: ranges from 0.6 to 15, with a mean of 4.64.
##  verbal: ranges from 1 to 10, with a mean of 6.66.
## gamble: ranges from 0 to 156, with a mean of 19.3.


## (B) Fit a linear model using the lm() function with gample variable as response, and the income variable as predictors, and report the regression coefficients

```{r} 
    model <- lm(gamble ~ income, data = teengamb)
    summary(model)$coefficients

```


## (C) Write the mathematical form of the model you fit in part (B). Clearly define each component in your model.

## The fitted model: y hat = beta 0 + beta 1 * income 
## Beta 0 is the intercept 
## Beta 1 is the slope 

## The regression coefficients are :
  
## The intercept is -6.324559
## And the slope coefficient is 5.520485
## The required regression equation is y hat = -6.324559 + 5.520485 * income 

##(D) Further numerical investigation: compute the mean and standard deviation of gamble and income for males (sex=0) and females (sex = 1) separately. Comment on the results.

``` {r}
    male_dataset <- teengamb[teengamb$sex == 0, ]
    female_dataset <- teengamb[teengamb$sex == 1, ]

    mean_sd <- data.frame(
      combined_dataset = c("Males", "Females"),
      mean_gamble = c(mean(male_dataset$gamble), mean(female_dataset$gamble)),
      sd_gamble = c(sd(male_dataset$gamble), sd(female_dataset$gamble)),
      mean_income = c(mean(male_dataset$income), mean(female_dataset$income)),
      sd_income = c(sd(male_dataset$income), sd(female_dataset$income))
    )
    print(mean_sd)


```


## Gambling Behavior:
## Males have a significantly higher average gambling expense (29.78) compared to females (3.87)
## Males have a significantly higher Gambling variability (SD) (37.32) than for females (5.15).
## Income:
## Males have a slightly higher average income (4.98) compared to females (4.15).
## Income variability is also greater for males (4.09) than for females (2.60).
## Males show both higher and more variable gambling expense and income compared to females.



## (E) Fit the same linear regression as in part (B), but separately for male and females. Report the regression coefficients.

``` {r}
    
    male_dataset <- teengamb[teengamb$sex == 0, ]
    female_dataset <- teengamb[teengamb$sex == 1, ]
    
    male_model <- lm(gamble ~ income, data = male_dataset)
    female_model <- lm(gamble ~ income, data = female_dataset)

    summary(male_model)
    summary(female_model)

```

## The regression coefficients for males
## The intercept is -2.660
## And the slope coefficient is 6.518
## The required regression equation for male is y hat =  -2.660 + 6.518 * income 

## The regression coefficients for females
## The intercept is 
## And the slope coefficient is .
## The required regression equation for female is, y hat =  3.1400 + 0.1749 * income 


## (F) Create a scatterplot between gamble (in y axis) and income (x axis), and color the points by sex. Then add two fitted regression lines from part (E) to the plot. Comment on the results.

``` {r}
    plot(teengamb$income, teengamb$gamble, col = teengamb$sex + 1)
    abline(male_model, col = "blue")
    abline(female_model, col = "red")
```

## If there is an upward trend rising from left-hand corner to upper right-hand corner, the correlation also looks like positive. If there is a downward trend, the correlation is negative.
## The males data show a strong positive correlation in the gambling expense as the income also increases. 
## While females are little bit consistent and show less gambling behavior, irrespective of income.