---
title: "SGupta_HW03Question1"
format: pdf
editor: visual
---

## Question 1

# (a) Which variables are statistically significant?

```{r}
library(faraway)
data(teengamb)
head(teengamb)

# Fit the full model
model <- lm(gamble ~ sex + status + income + verbal, data = teengamb)
summary(model)
```

Considering α = 0.05, sex is statistically significant as the p value 0.01 is less than 0.05.

Status is insignificant is this model preparation as the p value 0.8 is higher than alpha level.

income is highly significant. its p value is only 0.000017

And verbal is also not significant.

# (b) What interpretation should be given to the coefficient for sex?

```{r}
summary(model)$coefficients["sex", c("Estimate", "Std. Error","t value", "Pr(>|t|)")]
```

A unit increase in sex variable is associated with an average decrease in response variable gamble by about 22.12 units. A female variable tends to gamble less than male by 22.12 units.

# (c) Fit a model with just income as a predictor and use an F-test to compare it to the full model.

```{r}
model_income <- lm(gamble ~ income, data = teengamb)

anova(model_income, model)

```

-   The F-statistic is about 4.133 with 3 and 42 degrees of freedom.

-   The associated p-value is approximately 0.01177

-   The RSS difference between the two models is 6,384.8. This is due to the extra variables (variations) by adding the predictors sex, status, and verbal in the full model.

-   Since the p-value (0.01177) is less than the significance level (0.05), we reject the null hypothesis that the extra predictors (sex, status, and verbal) contribute no additional significance over the income-only model.
