---
title: "SGupta_HW03Question3"
format: pdf
editor: visual
---

## Question 3

# Fit a model with lpsa as the response and l cavol as the predictor.

```{r}
library(faraway)
set.seed(1001)
data(prostate)
names(prostate) 
model1 <- lm(lpsa ~ lcavol, data = prostate)
summary1 <- summary(model1)
summary1
```

# Record the residual standard error and the R2.

```{r}

rse <- summary(model1)$sigma
r2 <- summary(model1)$r.squared

print(paste("Residual Standard Error:", rse))
print(paste("Multiple R-squared:", r2))


```

# Now add lweight, svi, lpph, age, lcp, pgg45 and gleason to the model one at a time

```{r}
model2 <- lm(lpsa ~ lcavol + lweight, data = prostate)
summary2 <- summary(model2)

model3 <- lm(lpsa ~ lcavol + lweight + svi, data = prostate)
summary3 <- summary(model3)

model4 <- lm(lpsa ~ lcavol + lweight + svi + lbph, data = prostate)
summary4 <- summary(model4)

model5 <- lm(lpsa ~ lcavol + lweight + svi + lbph + age, data = prostate)
summary5 <- summary(model5)

model6 <- lm(lpsa ~ lcavol + lweight + svi + lbph + age + lcp, data = prostate)
summary6 <- summary(model6)

model7 <- lm(lpsa ~ lcavol + lweight + svi + lbph + age + lcp + pgg45, data = prostate)
summary7 <- summary(model7)

model8 <- lm(lpsa ~ lcavol + lweight + svi + lbph + age + lcp + pgg45 + gleason, data = prostate)
summary8 <- summary(model8)


```

```{r}
rse <- c(s1 = summary1$sigma,
         s2 = summary2$sigma,
         s3 = summary3$sigma,
         s4 = summary4$sigma,
         s5 = summary5$sigma,
         s6 = summary6$sigma,
         s7 = summary7$sigma,
         s8 = summary8$sigma)

r2 <- c(s1 = summary1$r.squared,
        s2 = summary2$r.squared,
        s3 = summary3$r.squared,
        s4 = summary4$r.squared,
        s5 = summary5$r.squared,
        s6 = summary6$r.squared,
        s7 = summary7$r.squared,
        s8 = summary8$r.squared)

rse
r2

```


# Plot the trends in these two statistics

```{r}
models <- 1:8
plot(models, rse, type="b", pch=19, col="blue",
     xlab="Model with number of predictors added)", ylab="Residual Standard Error",
     main="Trend in RSE as Predictors added in model")

plot(models, r2, type="b", pch=19, col="red",
     xlab="Model with number of predictors added)", ylab="R Squared",
     main="Trend in R2 as Predictors added in model")


```

Residual Standard Error (RSE): The RSE generally decreases as more predictors are added.
A lower RSE indicates that the model’s predictions are closer to the actual observed values. 

R square : The R square generally increases as more predictors are added, because R square can never decrease when more terms are added to a linear regression model.
The increase in R square shows that the added predictors explain additional variability in the response. 
After a certain point, each predictor addition only slightly affect R square. That means most of the significant variation is captures in the early predictors. 



